Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-05 p2/works-inverse.tex
Aligned proofs
Introducing valueless constructs

\begin{mydef}[Sets of sequences]
For brevity, we will use calligraphic letters (e.g. $\seqset{A, B, C}$ and $\seqset{S}$)
to denote sets of sequences,
and use $A\cc\seqset{B}$ to denote the set of sequences $\{A\cc B|B\in\seqset{B}\}$.
\end{mydef}

@@

\begin{mycor}
% An inference rule in the algebra
$B\eqext A \Rightarrow \worksc{A}{B}$.
\end{mycor}

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS:{}&\cxynv\indep \czwmv \\
&\quad\wedge \cxynv\aFS\neq\fsbroken \\
&\quad\wedge \czwmv\aFS\neq\fsbroken \\
&\Rightarrow \cxynv\cc \czwmv\aFS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
We proceed in an indirect way and
assume that $\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxynv\aFS)(m)\in\setvx{W}$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\noindent
We continue by defining inverse commands and sequences
which allow us to move parts of sequences between the
condition and consequence parts of $\workssign$.

@@

\begin{mydef}[$\typeeq$]
For two filesystems we write $\FS\typeeq\GS$
iff 
at all nodes the types of their values are the same, or they are both broken.
\end{mydef}

>>>>

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\seqset{A, B, C}$ and $\seqset{S}$)
to denote such sets for brevity.
We will use $A\cc\seqset{B}$ to denote the set of sequences $\{A\cc B|B\in\seqset{B}\}$.
\end{mydef}

@@

% \begin{mycor}
% % An inference rule in the algebra
% $B\eqext A \Rightarrow \worksc{A}{B}$.
% \end{mycor}

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\[ \cxynv\indep \czwmv \Rightarrow \worksc{\cxynv\cc \czwmv}{\cxynv, \czwmv}. \]
% \begin{align*}
% \forall\FS:{}&\cxynv\indep \czwmv \\
% &\quad\wedge \cxynv\aFS\neq\fsbroken \\
% &\quad\wedge \czwmv\aFS\neq\fsbroken \\
% &\Rightarrow \cxynv\cc \czwmv\aFS\neq\fsbroken.
% \end{align*}
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
We proceed in an indirect way and
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxynv\aFS)(m)\in\setvx{W}$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems we write $\FS\typeeq\GS$ iff they share the same nodes,
and at all nodes the types of their values are the same, or they are both broken.
\item For two commands we write $\alpha\typeeq\beta$ iff their input and output values,
and the node they operate on are the same.
\item For two sequences of commands we write $A\typeeq B$ iff they contain the same number of commands,
and pairwise the commands $\alpha\in A$ and $\beta\in B$ at the same indices are equivalent by type ($\alpha\typeeq\beta$).
\end{itemize}
\end{mydef}

This relation gives rise to equivalence classes of filesystems, commands, and sequences of commands,
each of which is associated with a new type of filesystem, command or sequence of commands in which
only types are noted, but not values.
(This is also equivalent to assuming $|\setf|=1$.)
We will call these constructs valueless.

\begin{mydef}[Valueless constructs]
$ $
\begin{itemize}
\item For any filesystem $\FS$, we write $\FS\T$ to denote the valueless filesystem assiociated with
the equivalenve class of $\FS$ by $\typeeq$; that is, where at each node $n$, $\FS\T$ contains the type of
$\FS(n)$. In general, a valueless filesystem is one where each node
is associated with a type, but not a value:
\begin{gather*}
\FS\T =
\begin{cases}
\setn \rightarrow \typeset \\
\fsbroken
\end{cases}
\end{gather*}

\item Similarly, for any command $\cxynv$ we write $\cxynv\T$ to denote the valueless command
$\cTxyn$, represented by a triple of the input type, output type and node.
The command $\cbrk$ has a separate valueless version, $\cbrk\T$.

\item Also, for any sequence of commands $S$ we write $S\T$ to denote the sequence of
valueless commands associated with the commands in $S$ in the same order; and for a set of sequences $\seqset{S}$
we write $\seqset{S}\T$ for $\{S\T|S\in\seqset{S}\}$.

\end{itemize}
\end{mydef}

A valueless command can be applied to a valueless filesystem with the expected outcome
that $\cTxyn$ will replace the type at $n$ with $Y$ 
if the original type was $X$ and the tree property still holds after the replacement.
This means that
the mapping $\T$ can be extended further to the 
the relations $\equiv$, $\eqext$ and $\workssign$, where
${\equiv}\T$, ${\eqext}\T$ and $\workssign\T$ hold or do not hold for sequences of valueless commands
depending on their behavior on valueless filesystems.

It is easy to see that 
$A\equiv B \Rightarrow A\T{\equiv}\T B\T$ 
and
$A\eqext B \Rightarrow A\T{\eqext}\T B\T$,
but the reverse is not true as 
$A\T$ and $B\T$ will have the same effect on all $\FS\T$
if $A$ and $B$ differ only in some of the values in their commands,
whereas clearly $A\nequiv B$.

With $\workssign$, however, the situation is different, as it only
depends on whether sequences break filesystems or not. Therefore the followings hold:
\begin{mycor}
\begin{gather*}
\works{\seqset{A}} \Leftrightarrow \worksT{\seqset{A}\T} \\
\worksc{\seqset{A}}{\seqset{B}} \Leftrightarrow \workscT{\seqset{A}\T}{\seqset{B}\T}
\end{gather*}
\end{mycor}

% ==================================================

We continue by defining inverse commands and sequences
which allow us to move parts of sequences between the
condition and consequence parts of $\workssign$.

@@
